---
title: "BRSM Data Processing and EDA"
author: "BRSM Project"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    fig_caption: true
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Create directory for output plots if it doesn't exist
if (!dir.exists("output_plots")) {
  dir.create("output_plots")
}

# Helper function for professional p-value reporting
format_pval <- function(p, threshold = 0.001) {
  if (is.na(p)) return(NA_character_)
  if (p < threshold) {
    return(paste0("< ", threshold))
  } else {
    return(sprintf("%.4f", p))
  }
}

# Safe column fetch for data sources with variant names
get_first_col <- function(df, candidates) {
  for (col_name in candidates) {
    if (col_name %in% names(df)) {
      return(df[[col_name]])
    }
  }
  rep(NA, nrow(df))
}
```

# PART I: DEMOGRAPHIC DATA PIPELINE

## 1. Constraint Definition

Missing values in the demographic dataset are filled only for valid participant entries.

Rules:

-   A row is valid only if participant_id (subId) is present.
-   Rows where participant_id is missing or empty are removed before imputation.

For data safety, this analysis uses the copy file:

-   Demographic data.xlsx (Local project file)

## 2. Load Data

```{r load-data}
library(tidyverse)
library(readxl)
library(writexl)
library(scales)

input_file <- "/home/prateek-tiwari/BRSM_Project/datafile /Demographic data.xlsx"

if (str_detect(str_to_lower(input_file), "\\.csv$")) {
  demo_data_raw <- read_csv(input_file, show_col_types = FALSE)
} else if (str_detect(str_to_lower(input_file), "\\.xlsx$")) {
  demo_data_raw <- read_excel(input_file)
} else {
  stop("Unsupported input format. Use .csv or .xlsx")
}

cat("Rows loaded:", nrow(demo_data_raw), "\n")
cat("Columns loaded:", ncol(demo_data_raw), "\n")
print(names(demo_data_raw))
```

## 3. Remove Invalid Rows (Critical Step)

```{r build-and-filter-participant-id}
# Build participant_id if not already present.
# If participant_id already exists, keep it.
if (!"participant_id" %in% names(demo_data_raw)) {
  demo_data <- demo_data_raw %>%
    mutate(participant_id = str_to_lower(str_extract(`Sub ID`, "(?i)sub\\d+")))
} else {
  demo_data <- demo_data_raw %>%
    mutate(participant_id = str_to_lower(participant_id))
}

rows_before_filter <- nrow(demo_data)

demo_data <- demo_data %>%
  filter(!is.na(participant_id) & participant_id != "")

rows_after_filter <- nrow(demo_data)

cat("Rows before participant_id filter:", rows_before_filter, "\n")
cat("Rows after participant_id filter:", rows_after_filter, "\n")
cat("Rows removed due to invalid participant_id:", rows_before_filter - rows_after_filter, "\n")
```

## 3.1 Condition Synchronization and Correction (Filename Ground Truth)

```{r condition-sync}
# Get paths for CSV files
participant_data_dir_sync <- if (dir.exists(file.path("/home/prateek-tiwari/BRSM_Project", "BRSM data csv"))) {
  file.path("/home/prateek-tiwari/BRSM_Project", "BRSM data csv")
} else {
  "/home/prateek-tiwari/BRSM_Project/BRSM data csv"
}
sync_files <- list.files(participant_data_dir_sync, pattern = "\\.csv$", full.names = TRUE)

# Create a lookup for conditions from filenames (Ground Truth)
filename_sync_lookup <- tibble(filepath = sync_files) %>%
  mutate(
    participant_id = str_to_lower(str_extract(basename(filepath), "(?i)sub\\d+")),
    # Extract AB or NB from filename
    filename_cond = str_to_upper(str_trim(str_extract(basename(filepath), "(?i)(?<=_)(AB|NB)\\s*(?=_)")))
  ) %>%
  filter(!is.na(participant_id), !is.na(filename_cond)) %>%
  distinct(participant_id, .keep_all = TRUE)

# Extract conditions recorded in Excel (Sub ID column)
demo_data <- demo_data %>%
  mutate(
    excel_cond = str_to_upper(str_trim(str_extract(`Sub ID`, "(?i)_(AB|NB)\\b"))) %>% str_remove("^_")
  )

# Join and detect mismatches
demo_data <- demo_data %>%
  left_join(filename_sync_lookup %>% select(participant_id, filename_cond), by = "participant_id")

# Log corrections explicitly
mismatches <- demo_data %>%
  filter(!is.na(filename_cond), !is.na(excel_cond), filename_cond != excel_cond)

if (nrow(mismatches) > 0) {
  cat("\n--- CONDITION CORRECTIONS (Based on Filenames) ---\n")
  for (i in 1:nrow(mismatches)) {
    cat(sprintf("Updated %s: Excel (%s) -> Filename (%s)\n",
                mismatches$participant_id[i], mismatches$excel_cond[i], mismatches$filename_cond[i]))
  }
  cat("-----------------------------------------------\n")
}

# Apply Ground Truth: Use Filename cond if available, otherwise fallback to Excel
demo_data <- demo_data %>%
  mutate(
    Condition = ifelse(!is.na(filename_cond), filename_cond, excel_cond)
  ) %>%
  select(-filename_cond, -excel_cond)
```

## 4. Inspect Missing Values (Before Imputation)

```{r missing-before}
colSums(is.na(demo_data))
```

## 5. Define Mode Function

```{r mode-function}
get_mode <- function(x) {
  x <- x[!is.na(x) & trimws(as.character(x)) != ""]
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
```

## 6. Compute Replacement Values

```{r replacement-values}
age_median <- median(as.numeric(demo_data$Age), na.rm = TRUE)

gender_mode <- get_mode(demo_data$Gender)
hand_mode   <- get_mode(demo_data$Handedness)
vision_mode <- get_mode(demo_data$Vision)

cat("Age median:", age_median, "\n")
cat("Gender mode:", gender_mode, "\n")
cat("Handedness mode:", hand_mode, "\n")
cat("Vision mode:", vision_mode, "\n")
```

## 7. Apply Imputation (Only Valid Rows)

```{r apply-imputation}
demo_clean <- demo_data %>%
  mutate(
    Age = as.numeric(Age),
    Age = ifelse(is.na(Age), age_median, Age),
    Gender = ifelse(is.na(Gender) | trimws(Gender) == "", gender_mode, Gender),
    Handedness = ifelse(is.na(Handedness) | trimws(Handedness) == "", hand_mode, Handedness),
    Vision = ifelse(is.na(Vision) | trimws(Vision) == "", vision_mode, Vision)
  )
```

## 8. Verification

```{r verification}
colSums(is.na(demo_clean))
```

Expected result: no missing values in Age, Gender, Handedness, Vision for valid participant rows.

## 9. Save Clean Data

```{r save-clean-data}
if (str_detect(str_to_lower(input_file), "\\.csv$")) {
  write_csv(demo_clean, input_file)
} else if (str_detect(str_to_lower(input_file), "\\.xlsx$")) {
  write_xlsx(demo_clean, input_file)
} else {
  stop("Unsupported input format. Use .csv or .xlsx")
}

cat("Updated file in place:", input_file, "\n")
```

## 10. Summary

```{r summary}
summary_tbl <- tibble(
  Metric = c(
    "Rows loaded",
    "Rows after participant_id validation",
    "Rows removed for missing/empty participant_id",
    "Missing Age after imputation",
    "Missing Gender after imputation",
    "Missing Handedness after imputation",
    "Missing Vision after imputation"
  ),
  Value = c(
    rows_before_filter,
    rows_after_filter,
    rows_before_filter - rows_after_filter,
    sum(is.na(demo_clean$Age)),
    sum(is.na(demo_clean$Gender) | trimws(demo_clean$Gender) == ""),
    sum(is.na(demo_clean$Handedness) | trimws(demo_clean$Handedness) == ""),
    sum(is.na(demo_clean$Vision) | trimws(demo_clean$Vision) == "")
  )
)

summary_tbl
```

## 10B. Participant Demographics Summary (Demographic data.xlsx)

```{r participant-demographics-summary}
# Build a clean demographic summary table similar to the report style shown.

id_source <- if ("Sub ID" %in% names(demo_data_raw)) {
  as.character(demo_data_raw$`Sub ID`)
} else if ("participant_id" %in% names(demo_data_raw)) {
  as.character(demo_data_raw$participant_id)
} else {
  rep(NA_character_, nrow(demo_data_raw))
}

demo_for_summary <- demo_data_raw %>%
  select(-any_of("Condition")) %>%
  mutate(
    id_source = id_source,
    participant_id = str_to_lower(str_extract(id_source, "(?i)sub\\d+")),
    Age_num = suppressWarnings(as.numeric(Age)),
    Gender_chr = as.character(Gender),
    Handedness_chr = as.character(Handedness),
    Vision_chr = as.character(Vision)
  ) %>%
  left_join(demo_data %>% select(participant_id, Condition), by = "participant_id") %>%
  filter(!is.na(participant_id), participant_id != "")

demo_complete <- demo_for_summary %>%
  filter(
    !is.na(Age_num),
    !is.na(Gender_chr) & trimws(Gender_chr) != "",
    !is.na(Handedness_chr) & trimws(Handedness_chr) != "",
    !is.na(Vision_chr) & trimws(Vision_chr) != ""
  ) %>%
  mutate(
    Gender_std = case_when(
      str_squish(str_to_lower(Gender_chr)) %in% c("female", "f") ~ "female",
      str_squish(str_to_lower(Gender_chr)) %in% c("male", "m") ~ "male",
      str_detect(str_squish(str_to_lower(Gender_chr)), "^female\\b") ~ "female",
      str_detect(str_squish(str_to_lower(Gender_chr)), "^male\\b") ~ "male",
      TRUE ~ "other"
    ),
    Handedness_std = case_when(
      str_squish(str_to_lower(Handedness_chr)) %in% c("right", "r", "right-handed", "right handed") ~ "right",
      str_squish(str_to_lower(Handedness_chr)) %in% c("left", "l", "left-handed", "left handed") ~ "left",
      str_detect(str_squish(str_to_lower(Handedness_chr)), "^right") ~ "right",
      str_detect(str_squish(str_to_lower(Handedness_chr)), "^left") ~ "left",
      TRUE ~ "other"
    ),
    Vision_std = case_when(
      str_detect(str_squish(str_to_lower(Vision_chr)), "uncorrected") ~ "uncorrected",
      str_detect(str_squish(str_to_lower(Vision_chr)), "corrected") ~ "corrected",
      str_squish(str_to_lower(Vision_chr)) %in% c("normal", "normal vision") ~ "normal",
      str_detect(str_squish(str_to_lower(Vision_chr)), "normal") ~ "normal",
      TRUE ~ "other"
    )
  )

demo_summary_tbl <- tibble(
  Characteristic = c(
    "N (complete demographic records)",
    "Age - Mean (SD)",
    "Age - Range",
    "Female",
    "Male",
    "Right-handed",
    "Left-handed",
    "Normal vision",
    "Corrected to normal",
    "Uncorrected vision difficulty"
  ),
  Value = c(
    nrow(demo_complete),
    sprintf("%.1f (%.1f)", mean(demo_complete$Age_num, na.rm = TRUE), sd(demo_complete$Age_num, na.rm = TRUE)),
    sprintf("%d - %d", floor(min(demo_complete$Age_num, na.rm = TRUE)), ceiling(max(demo_complete$Age_num, na.rm = TRUE))),
    sum(demo_complete$Gender_std == "female", na.rm = TRUE),
    sum(demo_complete$Gender_std == "male", na.rm = TRUE),
    sum(demo_complete$Handedness_std == "right", na.rm = TRUE),
    sum(demo_complete$Handedness_std == "left", na.rm = TRUE),
    sum(demo_complete$Vision_std == "normal", na.rm = TRUE),
    sum(demo_complete$Vision_std == "corrected", na.rm = TRUE),
    sum(demo_complete$Vision_std == "uncorrected", na.rm = TRUE)
  )
)

condition_demo_tbl <- demo_complete %>%
  filter(!is.na(Condition)) %>%
  group_by(Condition) %>%
  summarise(
    N = n(),
    `Age M` = round(mean(Age_num, na.rm = TRUE), 1),
    `Age SD` = round(sd(Age_num, na.rm = TRUE), 1),
    Female = sum(Gender_std == "female", na.rm = TRUE),
    Male = sum(Gender_std == "male", na.rm = TRUE),
    `Right-handed` = sum(Handedness_std == "right", na.rm = TRUE),
    `Left-handed` = sum(Handedness_std == "left", na.rm = TRUE),
    `Normal vision` = sum(Vision_std == "normal", na.rm = TRUE),
    `Corrected to normal` = sum(Vision_std == "corrected", na.rm = TRUE),
    `Uncorrected vision difficulty` = sum(Vision_std == "uncorrected", na.rm = TRUE),
    .groups = "drop"
  )

cat("Participant Demographics Summary\n")
cat("------------------------------\n")
cat("Total entries in Demographic data.xlsx:", nrow(demo_for_summary), "\n")
cat("Complete demographic records used for summary:", nrow(demo_complete), "\n\n")

demo_summary_tbl
condition_demo_tbl
```

## 10A. Data Preprocessing Report

```{r data-preprocessing-report}
participant_data_dir <- if (dir.exists(file.path("/home/prateek-tiwari/BRSM_Project", "BRSM data csv"))) {
  file.path("/home/prateek-tiwari/BRSM_Project", "BRSM data csv")
} else {
  "/home/prateek-tiwari/BRSM_Project/BRSM data csv"
}
participant_files_snapshot <- list.files(participant_data_dir, pattern = "\\.csv$", full.names = TRUE)
participant_ids_snapshot <- str_to_lower(str_extract(basename(participant_files_snapshot), "(?i)sub\\d+"))
participant_ids_snapshot <- participant_ids_snapshot[!is.na(participant_ids_snapshot)]

demo_ids_snapshot <- demo_clean %>%
  distinct(participant_id) %>%
  pull(participant_id)

matched_ids_snapshot <- intersect(participant_ids_snapshot, demo_ids_snapshot)
csv_not_in_demo_snapshot <- setdiff(participant_ids_snapshot, demo_ids_snapshot)
demo_not_in_csv_snapshot <- setdiff(demo_ids_snapshot, participant_ids_snapshot)

cat("Current Data Processing Snapshot\n")
cat("--------------------------------\n")
cat("Total participant CSV files found:", length(participant_files_snapshot), "\n")
cat("Unique participant IDs from CSV filenames:", length(unique(participant_ids_snapshot)), "\n")
cat("Rows in Demographic data.xlsx (before participant_id filter):", rows_before_filter, "\n")
cat("Rows in Demographic data.xlsx (after participant_id filter):", rows_after_filter, "\n")
cat("Unique participant IDs in cleaned demographics (Demographic data.xlsx):", length(unique(demo_ids_snapshot)), "\n")
cat("Participant IDs matched (CSV <-> demographics):", length(unique(matched_ids_snapshot)), "\n")
cat("CSV participant IDs missing in demographics:", length(unique(csv_not_in_demo_snapshot)), "\n")
cat("Demographic participant IDs missing in CSV:", length(unique(demo_not_in_csv_snapshot)), "\n")

if (length(demo_not_in_csv_snapshot) > 0) {
  cat("Example demographic IDs missing in CSV:", paste(head(demo_not_in_csv_snapshot, 10), collapse = ", "), "\n")
}

preprocessing_report <- tibble(
  Metric = c(
    "Total participant CSV files",
    "Unique participant IDs from CSV filenames",
    "Rows in Demographic data.xlsx before participant_id filter",
    "Rows in Demographic data.xlsx after participant_id filter",
    "Unique participant IDs in cleaned demographics",
    "Matched participant IDs (CSV vs demographics)",
    "CSV participant IDs missing in demographics",
    "Demographic participant IDs missing in CSV"
  ),
  Value = c(
    length(participant_files_snapshot),
    length(unique(participant_ids_snapshot)),
    rows_before_filter,
    rows_after_filter,
    length(unique(demo_ids_snapshot)),
    length(unique(matched_ids_snapshot)),
    length(unique(csv_not_in_demo_snapshot)),
    length(unique(demo_not_in_csv_snapshot))
  )
)

preprocessing_report
```

# PART II: BEHAVIORAL DATA PREPROCESSING & EXCLUSIONS

## 11. Fill Participant CSV Files by Matching Sub ID

```{r fill-participant-csvs-by-subid}
# Folder containing one CSV per participant
participant_data_dir <- if (dir.exists(file.path("/home/prateek-tiwari/BRSM_Project", "BRSM data csv"))) {
  file.path("/home/prateek-tiwari/BRSM_Project", "BRSM data csv")
} else {
  "/home/prateek-tiwari/BRSM_Project/BRSM data csv"
}
participant_files <- list.files(participant_data_dir, pattern = "\\.csv$", full.names = TRUE)

# Keep one cleaned demographic row per participant_id
demo_lookup <- demo_clean %>%
  distinct(participant_id, .keep_all = TRUE) %>%
  transmute(
    participant_id,
    age_ref = Age,
    gender_ref = Gender,
    handedness_ref = Handedness,
    vision_ref = Vision
  )

updated_files <- 0
matched_files <- 0
unmatched_files <- 0

for (f in participant_files) {
  # Match sub id from filename (example: sub17, sub150)
  subid <- str_to_lower(str_extract(basename(f), "(?i)sub\\d+"))

  if (is.na(subid) || !(subid %in% demo_lookup$participant_id)) {
    unmatched_files <- unmatched_files + 1
    next
  }

  ref <- demo_lookup %>% filter(participant_id == subid) %>% slice(1)

  dat <- read_csv(f, col_types = cols(.default = col_character()), show_col_types = FALSE)

  # Directly write demographic columns for this participant.
  # Other columns are left unchanged.
  if ("age" %in% names(dat)) {
    dat$age <- as.character(ref$age_ref)
  }
  if ("gender" %in% names(dat)) {
    dat$gender <- as.character(ref$gender_ref)
  }
  if ("handedness" %in% names(dat)) {
    dat$handedness <- as.character(ref$handedness_ref)
  }
  if ("vision" %in% names(dat)) {
    dat$vision <- as.character(ref$vision_ref)
  }

  write_csv(dat, f, na = "")
  matched_files <- matched_files + 1
  updated_files <- updated_files + 1
}

csv_fill_summary <- tibble(
  Metric = c(
    "Total participant CSV files found",
    "Files with matched sub ID in demographics",
    "Files updated in place",
    "Files without demographic sub ID match"
  ),
  Value = c(
    length(participant_files),
    matched_files,
    updated_files,
    unmatched_files
  )
)

csv_fill_summary
```

## 12. Exclusion Criterion

```{r exclusion-criterion}
# Criteria:
# 1) Exclude participants with encoding time > 27.05 minutes
# 2) Flag trials with RT < 0.2 s (200 ms)
# 3) Flag trials with RT > mean + 3SD

participant_data_dir <- if (dir.exists(file.path("/home/prateek-tiwari/BRSM_Project", "BRSM data csv"))) {
  file.path("/home/prateek-tiwari/BRSM_Project", "BRSM data csv")
} else {
  "/home/prateek-tiwari/BRSM_Project/BRSM data csv"
}
participant_files_excl <- list.files(participant_data_dir, pattern = "\\.csv$", full.names = TRUE)

parse_num <- function(x) suppressWarnings(readr::parse_number(as.character(x)))

encoding_records <- list()
trial_records <- list()

for (f in participant_files_excl) {
  pid <- str_to_lower(str_extract(basename(f), "(?i)sub\\d+"))
  dat <- read_csv(f, col_types = cols(.default = col_character()), show_col_types = FALSE)

  # Encoding time: recognition.started - instruction_2.stopped (in minutes)
  rec_start <- if ("recognition.started" %in% names(dat)) min(parse_num(dat$`recognition.started`), na.rm = TRUE) else NA_real_
  inst_stop <- if ("instruction_2.stopped" %in% names(dat)) min(parse_num(dat$`instruction_2.stopped`), na.rm = TRUE) else NA_real_

  if (is.infinite(rec_start)) rec_start <- NA_real_
  if (is.infinite(inst_stop)) inst_stop <- NA_real_

  encoding_min <- ifelse(!is.na(rec_start) & !is.na(inst_stop), (rec_start - inst_stop) / 60, NA_real_)

  encoding_records[[length(encoding_records) + 1]] <- tibble(
    participant_id = pid,
    encoding_minutes = encoding_min
  )

  # Trial-level RT extraction for criteria 2 and 3
  if (all(c("movie_id", "resp.rt") %in% names(dat))) {
    trial_tmp <- dat %>%
      transmute(
        participant_id = pid,
        movie_id_num = parse_num(movie_id),
        rt_sec = parse_num(`resp.rt`)
      ) %>%
      filter(!is.na(movie_id_num), movie_id_num >= 1, movie_id_num <= 40, !is.na(rt_sec))

    trial_records[[length(trial_records) + 1]] <- trial_tmp
  }
}

encoding_tbl <- if (length(encoding_records) > 0) {
  bind_rows(encoding_records) %>% distinct(participant_id, .keep_all = TRUE)
} else {
  tibble(participant_id = character(), encoding_minutes = numeric())
}

trial_tbl <- if (length(trial_records) > 0) {
  bind_rows(trial_records)
} else {
  tibble(participant_id = character(), movie_id_num = numeric(), rt_sec = numeric())
}

# Criterion 1
encoding_cutoff_min <- 27.05
excluded_participants_tbl <- encoding_tbl %>%
  filter(!is.na(encoding_minutes), encoding_minutes > encoding_cutoff_min)

# Criteria 2 and 3
rt_mean <- mean(trial_tbl$rt_sec, na.rm = TRUE)
rt_sd <- sd(trial_tbl$rt_sec, na.rm = TRUE)
rt_upper_cutoff <- rt_mean + 3 * rt_sd

rt_fast_tbl <- trial_tbl %>% filter(rt_sec < 0.2)
rt_slow_tbl <- trial_tbl %>% filter(rt_sec > rt_upper_cutoff)

cat("Exclusion / Flagging Summary\n")
cat("----------------------------\n")
cat("Total participant CSV files analyzed:", length(participant_files_excl), "\n")
cat("Participants with valid encoding time:", sum(!is.na(encoding_tbl$encoding_minutes)), "\n")
cat("Participants excluded (encoding > 27.05 min):", nrow(excluded_participants_tbl), "\n")
cat("Total trial rows evaluated for RT:", nrow(trial_tbl), "\n")
cat("Trials flagged as implausibly fast (RT < 0.2 s):", nrow(rt_fast_tbl), "\n")
cat("RT mean (s):", round(rt_mean, 4), "\n")
cat("RT SD (s):", round(rt_sd, 4), "\n")
cat("RT upper cutoff (mean + 3SD):", round(rt_upper_cutoff, 4), "\n")
cat("Trials flagged as slow outliers (RT > mean + 3SD):", nrow(rt_slow_tbl), "\n")

exclusion_report <- tibble(
  Criterion = c(
    "Participants excluded: encoding time > 27.05 minutes",
    "Trials flagged: RT < 0.2 s",
    "Trials flagged: RT > mean + 3SD"
  ),
  Count = c(
    nrow(excluded_participants_tbl),
    nrow(rt_fast_tbl),
    nrow(rt_slow_tbl)
  )
)

exclusion_report
```

## 13. Post-Exclusion Data Present in BRSM Data CSV

```{r post-exclusion-data-snapshot}
# Apply all three exclusion criteria to create the post-exclusion dataset:
# 1) Remove participants with encoding time > 27.05 min
# 2) Remove trials with RT < 0.2 s
# 3) Remove trials with RT > mean + 3SD

excluded_participant_ids <- excluded_participants_tbl$participant_id

post_exclusion_tbl <- trial_tbl %>%
  filter(!participant_id %in% excluded_participant_ids) %>%
  filter(rt_sec >= 0.2, rt_sec <= rt_upper_cutoff)

cat("Post-Exclusion Data Snapshot\n")
cat("----------------------------\n")
cat("Original trial rows:", nrow(trial_tbl), "\n")
cat("Trial rows after applying all exclusions:", nrow(post_exclusion_tbl), "\n")
cat("Unique participants after exclusion:", n_distinct(post_exclusion_tbl$participant_id), "\n")

post_exclusion_summary <- tibble(
  Metric = c(
    "Original trial rows",
    "Trial rows after all exclusions",
    "Unique participants after exclusion",
    "Min RT after exclusion (s)",
    "Max RT after exclusion (s)",
    "Mean RT after exclusion (s)"
  ),
  Value = c(
    nrow(trial_tbl),
    nrow(post_exclusion_tbl),
    n_distinct(post_exclusion_tbl$participant_id),
    round(min(post_exclusion_tbl$rt_sec, na.rm = TRUE), 4),
    round(max(post_exclusion_tbl$rt_sec, na.rm = TRUE), 4),
    round(mean(post_exclusion_tbl$rt_sec, na.rm = TRUE), 4)
  )
)

post_exclusion_summary

cat("\nFirst 10 rows of post-exclusion data:\n")
print(head(post_exclusion_tbl, 10))
```

Interpretation:

-   Rows without participant_id were removed first.
-   Missing values were filled only for valid participant rows.
-   Imputation used median for Age and mode for categorical variables.
-   Processing was performed on the copied file, not the original demographic source.
-   Participant CSV files were then updated in place only where sub ID matched cleaned demographics.

# PART III: DESCRIPTIVE STATISTICS & EDA

## 14. Corrected Participant Demographics Summary (Final)

```{r participant-demographics-summary-final}
# Recompute at the end so the final report always shows corrected demographic values.

id_source_final <- if ("Sub ID" %in% names(demo_data_raw)) {
  as.character(demo_data_raw$`Sub ID`)
} else if ("participant_id" %in% names(demo_data_raw)) {
  as.character(demo_data_raw$participant_id)
} else {
  rep(NA_character_, nrow(demo_data_raw))
}

demo_for_summary_final <- demo_data_raw %>%
  select(-any_of("Condition")) %>%
  mutate(
    id_source = id_source_final,
    participant_id = str_to_lower(str_extract(id_source, "(?i)sub\\d+")),
    Age_num = suppressWarnings(as.numeric(Age)),
    Gender_chr = as.character(Gender),
    Handedness_chr = as.character(Handedness),
    Vision_chr = as.character(Vision)
  ) %>%
  left_join(demo_data %>% select(participant_id, Condition), by = "participant_id") %>%
  filter(!is.na(participant_id), participant_id != "")

demo_complete_final <- demo_for_summary_final %>%
  filter(
    !is.na(Age_num),
    !is.na(Gender_chr) & trimws(Gender_chr) != "",
    !is.na(Handedness_chr) & trimws(Handedness_chr) != "",
    !is.na(Vision_chr) & trimws(Vision_chr) != ""
  ) %>%
  mutate(
    Gender_std = case_when(
      str_squish(str_to_lower(Gender_chr)) %in% c("female", "f") ~ "female",
      str_squish(str_to_lower(Gender_chr)) %in% c("male", "m") ~ "male",
      str_detect(str_squish(str_to_lower(Gender_chr)), "^female\\b") ~ "female",
      str_detect(str_squish(str_to_lower(Gender_chr)), "^male\\b") ~ "male",
      TRUE ~ "other"
    ),
    Handedness_std = case_when(
      str_squish(str_to_lower(Handedness_chr)) %in% c("right", "r", "right-handed", "right handed") ~ "right",
      str_squish(str_to_lower(Handedness_chr)) %in% c("left", "l", "left-handed", "left handed") ~ "left",
      str_detect(str_squish(str_to_lower(Handedness_chr)), "^right") ~ "right",
      str_detect(str_squish(str_to_lower(Handedness_chr)), "^left") ~ "left",
      TRUE ~ "other"
    ),
    Vision_std = case_when(
      str_detect(str_squish(str_to_lower(Vision_chr)), "uncorrected") ~ "uncorrected",
      str_detect(str_squish(str_to_lower(Vision_chr)), "corrected") ~ "corrected",
      str_squish(str_to_lower(Vision_chr)) %in% c("normal", "normal vision") ~ "normal",
      str_detect(str_squish(str_to_lower(Vision_chr)), "normal") ~ "normal",
      TRUE ~ "other"
    )
  )

demo_summary_tbl_final <- tibble(
  Characteristic = c(
    "N (complete demographic records)",
    "Age - Mean (SD)",
    "Age - Range",
    "Female",
    "Male",
    "Right-handed",
    "Left-handed",
    "Normal vision",
    "Corrected to normal",
    "Uncorrected vision difficulty"
  ),
  Value = c(
    nrow(demo_complete_final),
    sprintf("%.1f (%.1f)", mean(demo_complete_final$Age_num, na.rm = TRUE), sd(demo_complete_final$Age_num, na.rm = TRUE)),
    sprintf("%d - %d", floor(min(demo_complete_final$Age_num, na.rm = TRUE)), ceiling(max(demo_complete_final$Age_num, na.rm = TRUE))),
    sum(demo_complete_final$Gender_std == "female", na.rm = TRUE),
    sum(demo_complete_final$Gender_std == "male", na.rm = TRUE),
    sum(demo_complete_final$Handedness_std == "right", na.rm = TRUE),
    sum(demo_complete_final$Handedness_std == "left", na.rm = TRUE),
    sum(demo_complete_final$Vision_std == "normal", na.rm = TRUE),
    sum(demo_complete_final$Vision_std == "corrected", na.rm = TRUE),
    sum(demo_complete_final$Vision_std == "uncorrected", na.rm = TRUE)
  )
)

condition_demo_tbl_final <- demo_complete_final %>%
  filter(!is.na(Condition)) %>%
  group_by(Condition) %>%
  summarise(
    N = n(),
    `Age M` = round(mean(Age_num, na.rm = TRUE), 1),
    `Age SD` = round(sd(Age_num, na.rm = TRUE), 1),
    Female = sum(Gender_std == "female", na.rm = TRUE),
    Male = sum(Gender_std == "male", na.rm = TRUE),
    `Right-handed` = sum(Handedness_std == "right", na.rm = TRUE),
    `Left-handed` = sum(Handedness_std == "left", na.rm = TRUE),
    `Normal vision` = sum(Vision_std == "normal", na.rm = TRUE),
    `Corrected to normal` = sum(Vision_std == "corrected", na.rm = TRUE),
    `Uncorrected vision difficulty` = sum(Vision_std == "uncorrected", na.rm = TRUE),
    .groups = "drop"
  )

cat("Final Corrected Participant Demographics Summary\n")
cat("---------------------------------------------\n")
cat("Total entries in Demographic data.xlsx:", nrow(demo_for_summary_final), "\n")
cat("Complete demographic records used for summary:", nrow(demo_complete_final), "\n\n")

demo_summary_tbl_final
condition_demo_tbl_final
```

## 15. Descriptive Statistics (Updated From Current Data)

```{r build-descriptive-analysis-data, include=FALSE}
# Re-read participant CSV files so descriptive statistics reflect current updated data.
participant_data_dir_desc <- if (dir.exists(file.path("/home/prateek-tiwari/BRSM_Project", "BRSM data csv"))) {
  file.path("/home/prateek-tiwari/BRSM_Project", "BRSM data csv")
} else {
  "/home/prateek-tiwari/BRSM_Project/BRSM data csv"
}

participant_files_desc <- list.files(participant_data_dir_desc, pattern = "\\.csv$", full.names = TRUE)

read_desc_file <- function(filepath) {
  fname <- basename(filepath)
  pid <- str_to_lower(str_extract(fname, "(?i)sub\\d+"))

  dat <- read_csv(filepath, col_types = cols(.default = col_character()), show_col_types = FALSE)

  dat %>%
    mutate(participant_id = pid)
}

desc_raw <- map_dfr(participant_files_desc, read_desc_file)

# Integrate ground-truth Condition from demo_data
desc_trial <- desc_raw %>%
  inner_join(demo_data %>% select(participant_id, Condition), by = "participant_id") %>%
  mutate(
    movie_id_num = suppressWarnings(readr::parse_number(movie_id)),
    Accuracy = suppressWarnings(as.numeric(get_first_col(., c("resp.corr", "recogloop.resp.corr")))),
    RT = suppressWarnings(as.numeric(str_trim(str_remove_all(get_first_col(., c("resp.rt", "recogloop.resp.rt", "key_resp_3.rt")), "[\\[\\]\\s]")))),
    Confidence = suppressWarnings(as.numeric(get_first_col(., c("recogloop.conf_radio.response", "conf_radio.response", "confidence")))),
    FrameType = case_when(
      str_detect(target_img, "(?i)BB") ~ "BB",
      str_detect(target_img, "(?i)EM") ~ "EM",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(participant_id), !is.na(Condition)) %>%
  filter(!is.na(movie_id_num), movie_id_num >= 1, movie_id_num <= 40) %>%
  filter(!is.na(Accuracy), Accuracy %in% c(0, 1)) %>%
  filter(!participant_id %in% excluded_participants_tbl$participant_id) %>%
  filter(is.na(RT) | (RT >= 0.2 & RT <= rt_upper_cutoff)) %>%
  mutate(
    Condition = factor(Condition, levels = c("NB", "AB")),
    FrameType = factor(FrameType, levels = c("BB", "EM"))
  )

participant_level_desc <- desc_trial %>%
  group_by(participant_id, Condition) %>%
  summarise(
    mean_accuracy = mean(Accuracy, na.rm = TRUE),
    .groups = "drop"
  )

participant_frametype_desc <- desc_trial %>%
  filter(!is.na(FrameType)) %>%
  group_by(participant_id, Condition, FrameType) %>%
  summarise(
    mean_accuracy = mean(Accuracy, na.rm = TRUE),
    .groups = "drop"
  )
```

### 15.1 Overall Accuracy by Condition

```{r desc-overall-accuracy-by-condition}
overall_accuracy_by_condition <- participant_level_desc %>%
  group_by(Condition) %>%
  summarise(
    N = n(),
    Mean = round(mean(mean_accuracy, na.rm = TRUE), 3),
    SD = round(sd(mean_accuracy, na.rm = TRUE), 3),
    SE = round(SD / sqrt(N), 3),
    Min = round(min(mean_accuracy, na.rm = TRUE), 3),
    Max = round(max(mean_accuracy, na.rm = TRUE), 3),
    .groups = "drop"
  )

overall_accuracy_by_condition
```

### 15.2 Accuracy by Condition and Frame Type

```{r desc-accuracy-by-condition-frame-type}
accuracy_by_condition_frametype <- participant_frametype_desc %>%
  group_by(Condition, FrameType) %>%
  summarise(
    N = n(),
    Mean = round(mean(mean_accuracy, na.rm = TRUE), 3),
    SD = round(sd(mean_accuracy, na.rm = TRUE), 3),
    SE = round(SD / sqrt(N), 3),
    .groups = "drop"
  )

accuracy_by_condition_frametype
```

### 15.3 Reaction Time and Confidence by Condition

```{r desc-rt-confidence-by-condition}
rt_conf_by_condition <- desc_trial %>%
  group_by(Condition) %>%
  summarise(
    N = n(),
    Mean_RT = round(mean(RT, na.rm = TRUE), 2),
    SD_RT = round(sd(RT, na.rm = TRUE), 2),
    Mean_Conf = round(mean(Confidence, na.rm = TRUE), 2),
    SD_Conf = round(sd(Confidence, na.rm = TRUE), 2),
    .groups = "drop"
  )

rt_conf_by_condition
```

## 16. Exploratory Data Analysis (EDA)

This section provides visual and statistical insights into experimental performance, following the high-fidelity distribution and comparison styles used in formal reporting.

### 16.1 Distribution of Accuracy

Exploring how participant accuracy is distributed within and across conditions.

```{r eda-figures-1ab}
if (nrow(participant_level_desc) > 0 && any(!is.na(participant_level_desc$Condition))) {
  # Figure 1A: Histogram of Accuracy Faceted by Condition
  p1a <- ggplot(participant_level_desc, aes(x = mean_accuracy, fill = Condition)) +
    geom_histogram(bins = 12, color = "white", alpha = 0.8) +
    facet_wrap(~Condition, ncol = 1, labeller = label_both) +
    scale_x_continuous(labels = percent_format(), limits = c(0.4, 1.05)) +
    scale_fill_manual(values = c("NB" = "#3182bd", "AB" = "#e6550d")) +
    labs(title = "Histogram of Mean Recognition Accuracy per Participant",
         subtitle = "One value per participant (mean proportion correct across ~40 trials)",
         x = "Mean Accuracy (proportion correct)",
         y = "Number of Participants") +
    theme_classic() +
    theme(legend.position = "none", strip.background = element_rect(fill = "#f0f0f0"))
  p1a
  ggsave("output_plots/accuracy_histogram.png", p1a, width = 8, height = 6)

  # Figure 1B: Density Plot of Accuracy by Condition
  p1b <- ggplot(participant_level_desc, aes(x = mean_accuracy, fill = Condition, color = Condition)) +
    geom_density(alpha = 0.4, linewidth = 1) +
    scale_x_continuous(labels = percent_format(), limits = c(0.4, 1.05)) +
    scale_fill_manual(values = c("NB" = "#3182bd", "AB" = "#e6550d")) +
    scale_color_manual(values = c("NB" = "#3182bd", "AB" = "#e6550d")) +
    labs(title = "Density Plot of Mean Recognition Accuracy by Condition",
         subtitle = "Smooth probability distribution; NB (blue) vs AB (red)",
         x = "Mean Accuracy (proportion correct)",
         y = "Density") +
    theme_classic()
  p1b
  ggsave("output_plots/accuracy_density.png", p1b, width = 8, height = 6)
} else {
  cat("Skipping accuracy distribution plots: no valid participant data.\n")
}
```

### 16.2 Comparison by Condition

Directly comparing performance levels between Natural Blocking and Abrupt Blocking groups.

```{r eda-figure-2}
if (nrow(participant_level_desc) > 0 && any(!is.na(participant_level_desc$Condition))) {
  p2 <- ggplot(participant_level_desc, aes(x = Condition, y = mean_accuracy, fill = Condition)) +
    geom_violin(alpha = 0.5, color = "black") +
    geom_boxplot(width = 0.15, fill = "white", alpha = 0.5, outlier.shape = NA) +
    geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
    scale_y_continuous(labels = percent_format(), limits = c(0.4, 1.05)) +
    scale_fill_manual(values = c("NB" = "#3182bd", "AB" = "#e6550d")) +
    labs(title = "Violin Plot of Accuracy by Condition",
         subtitle = "Shape = full distribution; inner box = median and IQR; dots = individual participants",
         x = "Condition",
         y = "Mean Accuracy") +
    theme_classic() +
    theme(legend.position = "none")
  p2
  ggsave("output_plots/accuracy_condition_violin.png", p2, width = 8, height = 6)
} else {
  cat("Skipping condition comparison plot: no valid participant data.\n")
}
```

### 16.3 Comparison by Frame Type

Evaluation of performance based on stimulus frame types (Boundary vs. Middle).

```{r eda-figure-3a}
if (nrow(participant_frametype_desc) > 0 && any(!is.na(participant_frametype_desc$FrameType))) {
  p3a <- ggplot(participant_frametype_desc, aes(x = FrameType, y = mean_accuracy, fill = FrameType)) +
    geom_violin(alpha = 0.5, color = "black") +
    geom_boxplot(width = 0.15, fill = "white", alpha = 0.5, outlier.shape = NA) +
    geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
    scale_y_continuous(labels = percent_format(), limits = c(0.4, 1.05)) +
    scale_fill_manual(values = c("BB" = "#41ab5d", "EM" = "#feb24c")) +
    labs(title = "Violin Plot of Accuracy by Frame Type",
         subtitle = "BB = Boundary; EM = Event Middle; dots = individual participants",
         x = "Frame Type",
         y = "Mean Accuracy") +
    theme_classic() +
    theme(legend.position = "none")
  p3a
  ggsave("output_plots/accuracy_frametype_violin.png", p3a, width = 8, height = 6)
} else {
  cat("Skipping frame type plot: no valid frame type data.\n")
}
```

### 16.4 Interaction Heatmap

Visualizing the mean accuracy for each condition and frame type combination with exact values.

```{r eda-figure-3b}
if (nrow(participant_frametype_desc) > 0 && any(!is.na(participant_frametype_desc$Condition))) {
  heatmap_summary <- participant_frametype_desc %>%
    group_by(Condition, FrameType) %>%
    summarise(MeanAcc = mean(mean_accuracy, na.rm = TRUE), .groups = "drop")

  p3b <- ggplot(heatmap_summary, aes(x = FrameType, y = Condition, fill = MeanAcc)) +
    geom_tile(color = "white", linewidth = 2) +
    geom_text(aes(label = paste0(round(MeanAcc * 100, 1), "%")), color = "white", fontface = "bold", size = 8) +
    scale_fill_gradient(low = "#ef8a62", high = "#67a9cf", labels = percent_format()) +
    labs(title = "Heatmap of Mean Accuracy by Condition x Frame Type",
         subtitle = "Percentage labels indicate exact group means; Blue = higher accuracy",
         x = "Frame Type",
         y = "Condition",
         fill = "Mean Accuracy") +
    theme_classic() +
    theme(axis.line = element_blank())
  p3b
  ggsave("output_plots/accuracy_interaction_heatmap.png", p3b, width = 8, height = 6)
} else {
  cat("Skipping interaction heatmap: no valid frame type data.\n")
}
```

### 16.5 Consistency and Variability

Analysis of performance spread within groups.

```{r eda-variability-enhanced}
variability_summary <- participant_level_desc %>%
  group_by(Condition) %>%
  summarise(
    N = n(),
    SD = round(sd(mean_accuracy, na.rm = TRUE), 3),
    IQR = round(IQR(mean_accuracy, na.rm = TRUE), 3),
    Result_Range = paste0(round(min(mean_accuracy), 2), " - ", round(max(mean_accuracy), 2)),
    .groups = "drop"
  )

variability_summary
```

### 16.6 Interaction Performance with 95% CI

Detailed comparison of group means with 95% Confidence Intervals.

```{r eda-figure-4}
if (nrow(participant_frametype_desc) > 0 && any(!is.na(participant_frametype_desc$Condition))) {
  ci_summary <- participant_frametype_desc %>%
    group_by(Condition, FrameType) %>%
    summarise(
      N = n(),
      MeanAcc = mean(mean_accuracy, na.rm = TRUE),
      SD = sd(mean_accuracy, na.rm = TRUE),
      # Calculate 95% CI using t-distribution
      error = qt(0.975, df = N - 1) * (SD / sqrt(N)),
      lower = MeanAcc - error,
      upper = MeanAcc + error,
      .groups = "drop"
    )

  p4 <- ggplot(ci_summary, aes(x = FrameType, y = MeanAcc, fill = Condition)) +
    geom_bar(stat = "identity", position = position_dodge(0.8), width = 0.7, alpha = 0.8) +
    geom_errorbar(aes(ymin = lower, ymax = upper),
                  width = 0.2, position = position_dodge(0.8), color = "black") +
    scale_y_continuous(labels = percent_format(), limits = c(0, 1.05)) +
    scale_fill_manual(values = c("NB" = "#3182bd", "AB" = "#e6550d")) +
    labs(title = "Mean Accuracy (+/- 95% CI) by Condition and Frame Type",
         subtitle = "Error bars = 95% Confidence Interval (t-distribution, df = n - 1)",
         x = "Frame Type",
         y = "Mean Accuracy") +
    theme_classic()
  p4
  ggsave("output_plots/accuracy_interaction_ci.png", p4, width = 8, height = 6)
} else {
  cat("Skipping interaction CI plot: no valid frame type data.\n")
}
```

### 16.7 Age Distribution and Balance

Checking for age balance between conditions.

ggsave("output_plots/age_distribution.png", p5a, width = 8, height = 6)
ggsave("output_plots/age_condition_boxplot.png", p5b, width = 8, height = 6)
```{r eda-figures-5ab}
if (nrow(demo_complete_final) > 0 && any(!is.na(demo_complete_final$Condition))) {
  # Figure 5A: Age Histogram by Condition
  p5a <- ggplot(demo_complete_final, aes(x = Age_num, fill = Condition)) +
    geom_histogram(bins = 15, position = "dodge", alpha = 0.7) +
    scale_fill_manual(values = c("NB" = "#3182bd", "AB" = "#e6550d")) +
    labs(title = "Age Distribution by Condition",
         subtitle = "Each bar = number of participants of that age; NB (blue) vs AB (red)",
         x = "Age (years)",
         y = "Number of Participants") +
    theme_classic() +
    theme(legend.position = "right")
  p5a
  ggsave("output_plots/age_distribution.png", p5a, width = 8, height = 6)

  # Figure 5B: Age by Condition (Boxplot)
  p5b <- ggplot(demo_complete_final, aes(x = Condition, y = Age_num, fill = Condition)) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA) +
    geom_jitter(width = 0.1, alpha = 0.4, size = 1.5) +
    scale_fill_manual(values = c("NB" = "#3182bd", "AB" = "#e6550d")) +
    labs(title = "Age by Condition (Boxplot)",
         subtitle = "Median, IQR, and individual participants; NB vs AB",
         x = "Condition",
         y = "Age (years)") +
    theme_classic() +
    theme(legend.position = "none")
  p5b
  ggsave("output_plots/age_condition_boxplot.png", p5b, width = 8, height = 6)
} else {
  cat("Skipping age plots: no valid demographic data.\n")
}
```

### 16.8 Gender Breakdown and Composition

Analyzing the gender balance within the participant sample.

ggsave("output_plots/gender_breakdown.png", p5c, width = 8, height = 6)
gender_comp <- demo_complete_final %>%
ggsave("output_plots/gender_composition.png", p5d, width = 8, height = 6)
```{r eda-figures-5cd}
if (nrow(demo_complete_final) > 0 && any(!is.na(demo_complete_final$Gender_std))) {
  # Figure 5C: Overall Gender Breakdown
  gender_total <- demo_complete_final %>%
    group_by(Gender_std) %>%
    summarise(Count = n(), .groups = "drop") %>%
    mutate(Perc = Count / sum(Count),
           Label = paste0(Count, "\n(", percent(Perc, accuracy = 1), ")"),
           Gender_std = str_to_title(Gender_std))

  p5c <- ggplot(gender_total, aes(x = Gender_std, y = Count, fill = Gender_std)) +
    geom_bar(stat = "identity", width = 0.6) +
    geom_text(aes(label = Label), vjust = -0.5, fontface = "bold", size = 5) +
    scale_fill_manual(values = c("Female" = "#e7298a", "Male" = "#377eb8", "Other" = "#999999")) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    labs(title = "Gender Breakdown of Participants",
         subtitle = "Full sample (complete demographic records only)",
         x = "Gender",
         y = "Number of Participants") +
    theme_classic() +
    theme(legend.position = "none", axis.line.y = element_blank(), axis.ticks.y = element_blank())
  p5c
  ggsave("output_plots/gender_breakdown.png", p5c, width = 8, height = 6)

  # Figure 5D: Gender Composition by Condition
  gender_comp <- demo_complete_final %>%
    group_by(Condition, Gender_std) %>%
    summarise(Count = n(), .groups = "drop_last") %>%
    mutate(Perc = Count / sum(Count),
           Label = paste0(Count, "\n(", percent(Perc, accuracy = 1), ")"),
           Gender_std = str_to_title(Gender_std))

  p5d <- ggplot(gender_comp, aes(x = Condition, y = Perc, fill = Gender_std)) +
    geom_bar(stat = "identity", position = "fill", width = 0.7, alpha = 0.9) +
    geom_text(aes(label = Label), position = position_stack(vjust = 0.5), color = "white", fontface = "bold", size = 4) +
    scale_y_continuous(labels = percent_format()) +
    scale_fill_manual(values = c("Female" = "#e7298a", "Male" = "#377eb8", "Other" = "#999999")) +
    labs(title = "Gender Composition by Condition",
         subtitle = "Proportional stacked bar; numbers = count and percentage within condition",
         x = "Condition",
         y = "Proportion of Participants",
         fill = "Gender") +
    theme_classic()
  p5d
  ggsave("output_plots/gender_composition.png", p5d, width = 8, height = 6)
} else {
  cat("Skipping gender plots: no valid demographic data.\n")
}
```
