---
title: "BRSM Hypotheses, Tests, and Results"
author: "BRSM Project"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: true
    fig_caption: true
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

format_pval <- function(p, threshold = 0.001) {
  if (is.na(p)) return(NA_character_)
  if (p < threshold) {
    return(paste0("< ", threshold))
  }
  sprintf("%.4f", p)
}

get_first_col <- function(df, candidates) {
  for (col_name in candidates) {
    if (col_name %in% names(df)) {
      return(df[[col_name]])
    }
  }
  rep(NA, nrow(df))
}
```

# Data Inputs (Used for All Tests)

```{r load-packages}
library(tidyverse)
library(readxl)
library(scales)
library(ARTool)
library(coin)
library(lme4)
library(car)
```

```{r load-demographics}
input_file <- "/home/prateek-tiwari/BRSM_Project/datafile /Demographic data.xlsx"
participant_data_dir <- "/home/prateek-tiwari/BRSM_Project/BRSM data csv"

if (str_detect(str_to_lower(input_file), "\\.csv$")) {
  demo_data_raw <- read_csv(input_file, show_col_types = FALSE)
} else if (str_detect(str_to_lower(input_file), "\\.xlsx$")) {
  demo_data_raw <- read_excel(input_file)
} else {
  stop("Unsupported input format. Use .csv or .xlsx")
}

# Build participant_id
if (!"participant_id" %in% names(demo_data_raw)) {
  demo_data <- demo_data_raw %>%
    mutate(participant_id = str_to_lower(str_extract(`Sub ID`, "(?i)sub\\d+")))
} else {
  demo_data <- demo_data_raw %>%
    mutate(participant_id = str_to_lower(participant_id))
}

demo_data <- demo_data %>%
  filter(!is.na(participant_id) & participant_id != "")

# Condition sync from filenames
sync_files <- list.files(participant_data_dir, pattern = "\\.csv$", full.names = TRUE)
filename_sync_lookup <- tibble(filepath = sync_files) %>%
  mutate(
    participant_id = str_to_lower(str_extract(basename(filepath), "(?i)sub\\d+")),
    filename_cond = str_to_upper(str_trim(str_extract(basename(filepath), "(?i)(?<=_)(AB|NB)\\s*(?=_)")))
  ) %>%
  filter(!is.na(participant_id), !is.na(filename_cond)) %>%
  distinct(participant_id, .keep_all = TRUE)

demo_data <- demo_data %>%
  mutate(
    excel_cond = str_to_upper(str_trim(str_extract(`Sub ID`, "(?i)_(AB|NB)\\b"))) %>% str_remove("^_")
  ) %>%
  left_join(filename_sync_lookup %>% select(participant_id, filename_cond), by = "participant_id") %>%
  mutate(
    Condition = ifelse(!is.na(filename_cond), filename_cond, excel_cond)
  ) %>%
  select(-filename_cond, -excel_cond)
```

```{r exclusion-criterion}
parse_num <- function(x) suppressWarnings(readr::parse_number(as.character(x)))
participant_files_excl <- list.files(participant_data_dir, pattern = "\\.csv$", full.names = TRUE)

encoding_records <- list()
trial_records <- list()

for (f in participant_files_excl) {
  pid <- str_to_lower(str_extract(basename(f), "(?i)sub\\d+"))
  dat <- read_csv(f, col_types = cols(.default = col_character()), show_col_types = FALSE)

  rec_start <- if ("recognition.started" %in% names(dat)) min(parse_num(dat$`recognition.started`), na.rm = TRUE) else NA_real_
  inst_stop <- if ("instruction_2.stopped" %in% names(dat)) min(parse_num(dat$`instruction_2.stopped`), na.rm = TRUE) else NA_real_

  if (is.infinite(rec_start)) rec_start <- NA_real_
  if (is.infinite(inst_stop)) inst_stop <- NA_real_

  encoding_min <- ifelse(!is.na(rec_start) & !is.na(inst_stop), (rec_start - inst_stop) / 60, NA_real_)

  encoding_records[[length(encoding_records) + 1]] <- tibble(
    participant_id = pid,
    encoding_minutes = encoding_min
  )

  if (all(c("movie_id", "resp.rt") %in% names(dat))) {
    trial_tmp <- dat %>%
      transmute(
        participant_id = pid,
        movie_id_num = parse_num(movie_id),
        rt_sec = parse_num(`resp.rt`)
      ) %>%
      filter(!is.na(movie_id_num), movie_id_num >= 1, movie_id_num <= 40, !is.na(rt_sec))

    trial_records[[length(trial_records) + 1]] <- trial_tmp
  }
}

encoding_tbl <- if (length(encoding_records) > 0) {
  bind_rows(encoding_records) %>% distinct(participant_id, .keep_all = TRUE)
} else {
  tibble(participant_id = character(), encoding_minutes = numeric())
}

trial_tbl <- if (length(trial_records) > 0) {
  bind_rows(trial_records)
} else {
  tibble(participant_id = character(), movie_id_num = numeric(), rt_sec = numeric())
}

encoding_cutoff_min <- 27.05
excluded_participants_tbl <- encoding_tbl %>%
  filter(!is.na(encoding_minutes), encoding_minutes > encoding_cutoff_min)

rt_mean <- mean(trial_tbl$rt_sec, na.rm = TRUE)
rt_sd <- sd(trial_tbl$rt_sec, na.rm = TRUE)
rt_upper_cutoff <- rt_mean + 3 * rt_sd
```

```{r build-analysis-data}
participant_files_desc <- list.files(participant_data_dir, pattern = "\\.csv$", full.names = TRUE)

read_desc_file <- function(filepath) {
  fname <- basename(filepath)
  pid <- str_to_lower(str_extract(fname, "(?i)sub\\d+"))
  dat <- read_csv(filepath, col_types = cols(.default = col_character()), show_col_types = FALSE)
  dat %>% mutate(participant_id = pid)
}

desc_raw <- map_dfr(participant_files_desc, read_desc_file)

desc_trial <- desc_raw %>%
  inner_join(demo_data %>% select(participant_id, Condition), by = "participant_id") %>%
  mutate(
    movie_id_num = suppressWarnings(readr::parse_number(movie_id)),
    Accuracy = suppressWarnings(as.numeric(get_first_col(., c("resp.corr", "recogloop.resp.corr")))),
    RT = suppressWarnings(as.numeric(str_trim(str_remove_all(get_first_col(., c("resp.rt", "recogloop.resp.rt", "key_resp_3.rt")), "[\\[\\]\\s]")))),
    Confidence = suppressWarnings(as.numeric(get_first_col(., c("recogloop.conf_radio.response", "conf_radio.response", "confidence")))),
    FrameType = case_when(
      str_detect(target_img, "(?i)BB") ~ "BB",
      str_detect(target_img, "(?i)EM") ~ "EM",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(participant_id), !is.na(Condition)) %>%
  filter(!is.na(movie_id_num), movie_id_num >= 1, movie_id_num <= 40) %>%
  filter(!is.na(Accuracy), Accuracy %in% c(0, 1)) %>%
  filter(!participant_id %in% excluded_participants_tbl$participant_id) %>%
  filter(is.na(RT) | (RT >= 0.2 & RT <= rt_upper_cutoff)) %>%
  mutate(
    Condition = factor(Condition, levels = c("NB", "AB")),
    FrameType = factor(FrameType, levels = c("BB", "EM"))
  )

participant_level_desc <- desc_trial %>%
  group_by(participant_id, Condition) %>%
  summarise(mean_accuracy = mean(Accuracy, na.rm = TRUE), .groups = "drop")

participant_frametype_desc <- desc_trial %>%
  filter(!is.na(FrameType)) %>%
  group_by(participant_id, Condition, FrameType) %>%
  summarise(mean_accuracy = mean(Accuracy, na.rm = TRUE), .groups = "drop")
```

# Experimental Hypotheses H1–H8

## H1: Condition Effect (NB > AB Accuracy)

```{r h1-tests}
participant_level_desc <- participant_level_desc %>%
  mutate(Condition = factor(Condition, levels = c("NB", "AB")))

m_art <- art(mean_accuracy ~ Condition, data = participant_level_desc)
anova_art <- as.data.frame(anova(m_art))

mw_result <- wilcox.test(mean_accuracy ~ Condition, data = participant_level_desc, conf.int = TRUE)

perm_test <- oneway_test(mean_accuracy ~ Condition, data = participant_level_desc,
                         distribution = approximate(nresample = 10000))

overall_median <- median(participant_level_desc$mean_accuracy, na.rm = TRUE)
median_table <- table(
  participant_level_desc$Condition,
  participant_level_desc$mean_accuracy > overall_median
)
med_test <- chisq.test(median_table)

h1_summary <- tibble(
  Test = c("ART ANOVA", "Mann-Whitney U", "Permutation Test", "Median Test"),
  p_value = c(
    format_pval(anova_art[1, "Pr(>F)"]),
    format_pval(mw_result$p.value),
    format_pval(as.numeric(coin::pvalue(perm_test))),
    format_pval(med_test$p.value)
  )
)

h1_summary
```

## H2: Boundary Advantage (BB vs EM)

```{r h2-tests}
h2_wide <- participant_frametype_desc %>%
  select(participant_id, FrameType, mean_accuracy) %>%
  pivot_wider(names_from = FrameType, values_from = mean_accuracy) %>%
  filter(!is.na(BB), !is.na(EM))

participant_frametype_desc <- participant_frametype_desc %>%
  mutate(FrameType = factor(FrameType, levels = c("EM", "BB")))

m_h2_art <- art(mean_accuracy ~ FrameType + (1|participant_id), data = participant_frametype_desc)
anova_h2_art <- as.data.frame(anova(m_h2_art))

h2_wilcoxon_result <- wilcox.test(h2_wide$BB, h2_wide$EM, paired = TRUE, conf.int = TRUE)

h2_diffs <- h2_wide$BB - h2_wide$EM
pos_diffs <- sum(h2_diffs > 0, na.rm = TRUE)
total_valid <- sum(h2_diffs != 0, na.rm = TRUE)
h2_sign_result <- binom.test(pos_diffs, total_valid, p = 0.5, alternative = "two.sided")

h2_friedman_result <- friedman.test(as.matrix(h2_wide[, c("BB", "EM")]))

h2_summary_tbl <- tibble(
  Test = c("ART ANOVA", "Wilcoxon Signed-Rank", "Sign Test", "Friedman Test"),
  p_value = c(
    format_pval(anova_h2_art[1, "Pr(>F)"]),
    format_pval(h2_wilcoxon_result$p.value),
    format_pval(h2_sign_result$p.value),
    format_pval(h2_friedman_result$p.value)
  )
)

h2_summary_tbl
```

## H3: Event-Middle Stability (EM: NB vs AB)

```{r h3-tests}
h3_data <- participant_frametype_desc %>%
  filter(FrameType == "EM") %>%
  select(participant_id, Condition, mean_accuracy)

m_h3_art <- art(mean_accuracy ~ Condition, data = h3_data)
anova_h3_art <- as.data.frame(anova(m_h3_art))

h3_mw_result <- wilcox.test(mean_accuracy ~ Condition, data = h3_data, conf.int = TRUE)

h3_perm_test <- oneway_test(mean_accuracy ~ Condition, data = h3_data,
                            distribution = approximate(nresample = 10000))

set.seed(42)
nb_em_vals <- h3_data$mean_accuracy[h3_data$Condition == "NB"]
ab_em_vals <- h3_data$mean_accuracy[h3_data$Condition == "AB"]
boot_diffs <- replicate(10000, {
  mean(sample(nb_em_vals, replace = TRUE)) - mean(sample(ab_em_vals, replace = TRUE))
})

h3_ci_lower <- quantile(boot_diffs, 0.00625)
h3_ci_upper <- quantile(boot_diffs, 0.99375)

h3_summary_tbl <- tibble(
  Test = c("ART ANOVA", "Mann-Whitney U", "Permutation Test", "Bootstrap (CI includes 0)"),
  Result = c(
    format_pval(anova_h3_art[1, "Pr(>F)"]),
    format_pval(h3_mw_result$p.value),
    format_pval(as.numeric(coin::pvalue(h3_perm_test))),
    ifelse(h3_ci_lower < 0 & h3_ci_upper > 0, "Yes", "No")
  )
)

h3_summary_tbl
```

## H4: Interaction (Boundary Advantage Stronger in NB)

```{r h4-tests}
h4_data <- participant_frametype_desc %>%
  select(participant_id, Condition, FrameType, mean_accuracy) %>%
  pivot_wider(names_from = FrameType, values_from = mean_accuracy) %>%
  mutate(diff_score = BB - EM) %>%
  filter(!is.na(diff_score))

m_h4_art <- art(mean_accuracy ~ FrameType * Condition + (1|participant_id), data = participant_frametype_desc)
anova_h4_art <- as.data.frame(anova(m_h4_art))

h4_mw_result <- wilcox.test(diff_score ~ Condition, data = h4_data, conf.int = TRUE)

h4_perm_test <- oneway_test(diff_score ~ Condition, data = h4_data,
                            distribution = approximate(nresample = 10000))

h4_friedman_result <- friedman.test(as.matrix(h4_data[, c("BB", "EM")]))

h4_summary_tbl <- tibble(
  Test = c("ART ANOVA (Interaction)", "Mann-Whitney (Advantage)", "Permutation Test", "Friedman Test"),
  p_value = c(
    format_pval(anova_h4_art["FrameType:Condition", "Pr(>F)"]),
    format_pval(h4_mw_result$p.value),
    format_pval(as.numeric(coin::pvalue(h4_perm_test))),
    format_pval(h4_friedman_result$p.value)
  )
)

h4_summary_tbl
```

## H5: Confidence–Accuracy Relationship

```{r h5-tests}
h5_data <- desc_trial %>%
  filter(!is.na(Confidence), !is.na(Accuracy)) %>%
  select(participant_id, Condition, Accuracy, Confidence) %>%
  group_by(participant_id) %>%
  mutate(Conf_z = as.numeric(scale(Confidence))) %>%
  ungroup() %>%
  filter(!is.na(Conf_z))

nb_corr <- cor.test(h5_data$Confidence[h5_data$Condition == "NB"],
                    h5_data$Accuracy[h5_data$Condition == "NB"], method = "spearman")
ab_corr <- cor.test(h5_data$Confidence[h5_data$Condition == "AB"],
                    h5_data$Accuracy[h5_data$Condition == "AB"], method = "spearman")

h5_mw_conf <- wilcox.test(Confidence ~ Condition, data = h5_data)

m_h5_glmm <- glmer(Accuracy ~ Conf_z * Condition + (1 | participant_id),
                   data = h5_data, family = binomial)

sum_h5_glmm <- as.data.frame(coef(summary(m_h5_glmm)))
int_term <- grep("Conf_z:Condition", rownames(sum_h5_glmm), value = TRUE)
h5_int_p <- if(length(int_term) > 0) sum_h5_glmm[int_term, "Pr(>|z|)"] else NA_real_

set.seed(42)
obs_diff <- nb_corr$estimate - ab_corr$estimate
perm_diffs <- replicate(10000, {
  shuffled_cond <- sample(h5_data$Condition)
  cor(h5_data$Confidence[shuffled_cond == "NB"], h5_data$Accuracy[shuffled_cond == "NB"], method = "spearman") -
    cor(h5_data$Confidence[shuffled_cond == "AB"], h5_data$Accuracy[shuffled_cond == "AB"], method = "spearman")
})

h5_perm_p <- mean(abs(perm_diffs) >= abs(obs_diff))

h5_summary_tbl <- tibble(
  Test = c("Spearman (NB)", "Spearman (AB)", "Mann-Whitney", "Interaction GLMM", "Permutation"),
  Result = c(
    sprintf("rho = %.3f, p = %s", nb_corr$estimate, format_pval(nb_corr$p.value)),
    sprintf("rho = %.3f, p = %s", ab_corr$estimate, format_pval(ab_corr$p.value)),
    sprintf("p = %s", format_pval(h5_mw_conf$p.value)),
    sprintf("p = %s", format_pval(h5_int_p)),
    sprintf("p = %s", format_pval(h5_perm_p))
  )
)

h5_summary_tbl
```

## H6: Difficulty Moderation

```{r h6-tests}
stimulus_difficulty_map <- desc_trial %>%
  group_by(movie_id_num) %>%
  summarise(mean_acc = mean(Accuracy, na.rm = TRUE), .groups = "drop") %>%
  mutate(Difficulty = factor(ifelse(mean_acc < median(mean_acc), "Hard", "Easy")))

h6_trial_data <- desc_trial %>%
  inner_join(stimulus_difficulty_map %>% select(movie_id_num, Difficulty), by = "movie_id_num")

h6_diff_data <- h6_trial_data %>%
  group_by(participant_id, FrameType, Difficulty) %>%
  summarise(mean_acc = mean(Accuracy, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = FrameType, values_from = mean_acc) %>%
  mutate(diff_score = BB - EM) %>%
  filter(!is.na(diff_score))

h6_kw_result <- kruskal.test(diff_score ~ Difficulty, data = h6_diff_data)

m_h6_art <- art(Accuracy ~ FrameType * Difficulty + (1|participant_id), data = h6_trial_data)
anova_h6_art <- as.data.frame(anova(m_h6_art))

h6_mw_result <- wilcox.test(diff_score ~ Difficulty, data = h6_diff_data)

h6_perm_test <- oneway_test(diff_score ~ factor(Difficulty), data = h6_diff_data,
                            distribution = approximate(nresample = 10000))

h6_summary_tbl <- tibble(
  Test = c("Kruskal-Wallis", "ART ANOVA Interaction", "Mann-Whitney", "Permutation"),
  p_value = c(
    format_pval(h6_kw_result$p.value),
    format_pval(anova_h6_art["FrameType:Difficulty", "Pr(>F)"]),
    format_pval(h6_mw_result$p.value),
    format_pval(as.numeric(coin::pvalue(h6_perm_test)))
  )
)

h6_summary_tbl
```

## H7: Variability Effect (AB variance > NB variance)

```{r h7-tests}
h7_data <- participant_level_desc %>%
  filter(!is.na(mean_accuracy)) %>%
  select(participant_id, Condition, mean_accuracy)

h7_levene_result <- car::leveneTest(mean_accuracy ~ Condition, data = h7_data)
h7_bf_result <- car::leveneTest(mean_accuracy ~ Condition, data = h7_data, center = median)

set.seed(42)
sd_nb <- sd(h7_data$mean_accuracy[h7_data$Condition == "NB"])
sd_ab <- sd(h7_data$mean_accuracy[h7_data$Condition == "AB"])
obs_sd_diff <- sd_ab - sd_nb

perm_sd_diffs <- replicate(10000, {
  shuffled_cond <- sample(h7_data$Condition)
  sd(h7_data$mean_accuracy[shuffled_cond == "AB"]) - sd(h7_data$mean_accuracy[shuffled_cond == "NB"])
})

h7_perm_p <- mean(perm_sd_diffs >= obs_sd_diff)

boot_var_diffs <- replicate(10000, {
  var(sample(h7_data$mean_accuracy[h7_data$Condition == "AB"], replace = TRUE)) -
    var(sample(h7_data$mean_accuracy[h7_data$Condition == "NB"], replace = TRUE))
})

h7_ci_lower <- quantile(boot_var_diffs, 0.00625)
h7_ci_upper <- quantile(boot_var_diffs, 0.99375)

h7_summary_tbl <- tibble(
  Test = c("Levene", "Brown-Forsythe", "Permutation", "Bootstrap (CI > 0)"),
  Result = c(
    format_pval(h7_levene_result[["Pr(>F)"]][1]),
    format_pval(h7_bf_result[["Pr(>F)"]][1]),
    format_pval(h7_perm_p),
    ifelse(h7_ci_lower > 0, "Yes", "No")
  )
)

h7_summary_tbl
```

## H8: Encoding Strength Effect

```{r h8-tests}
h8_diff_data <- participant_frametype_desc %>%
  select(participant_id, FrameType, mean_accuracy) %>%
  pivot_wider(names_from = FrameType, values_from = mean_accuracy) %>%
  mutate(diff_score = BB - EM)

h8_data <- h8_diff_data %>%
  inner_join(encoding_tbl %>% select(participant_id, encoding_minutes), by = "participant_id") %>%
  filter(!is.na(diff_score), !is.na(encoding_minutes))

h8_spearman_result <- cor.test(h8_data$encoding_minutes, h8_data$diff_score, method = "spearman")

m_h8_reg <- lm(diff_score ~ encoding_minutes, data = h8_data)
sum_h8_reg <- summary(m_h8_reg)
h8_reg_p <- coef(sum_h8_reg)[2, "Pr(>|t|)"]

h8_data_split <- h8_data %>%
  mutate(Encoding_Group = ifelse(encoding_minutes > median(encoding_minutes), "Longer", "Shorter"))

h8_mw_result <- wilcox.test(diff_score ~ Encoding_Group, data = h8_data_split)

set.seed(42)
obs_corr <- cor(h8_data$encoding_minutes, h8_data$diff_score, method = "pearson")
perm_corrs <- replicate(10000, {
  cor(sample(h8_data$encoding_minutes), h8_data$diff_score, method = "pearson")
})

h8_perm_p <- mean(abs(perm_corrs) >= abs(obs_corr))

h8_summary_tbl <- tibble(
  Test = c("Spearman", "Linear Regression", "Median Split", "Permutation"),
  Result = c(
    sprintf("rho = %.3f", h8_spearman_result$estimate),
    sprintf("p = %s", format_pval(h8_reg_p)),
    sprintf("p = %s", format_pval(h8_mw_result$p.value)),
    sprintf("p = %s", format_pval(h8_perm_p))
  )
)

h8_summary_tbl
```

# Results Summary (Pass/Fail)

```{r results-summary}
# H1: primary decision from Mann-Whitney
h1_pass <- mw_result$p.value < 0.05

# H2: require >=2 significant tests and BB > EM
h2_pvals <- c(
  anova_h2_art[1, "Pr(>F)"],
  h2_wilcoxon_result$p.value,
  h2_sign_result$p.value,
  h2_friedman_result$p.value
)
h2_sig_count <- sum(h2_pvals < 0.0125, na.rm = TRUE)
bb_med <- median(h2_wide$BB, na.rm = TRUE)
em_med <- median(h2_wide$EM, na.rm = TRUE)
h2_pass <- h2_sig_count >= 2 && bb_med > em_med

# H3: stability supported by non-significance and CI including 0
h3_pvals <- c(
  anova_h3_art[1, "Pr(>F)"],
  h3_mw_result$p.value,
  as.numeric(coin::pvalue(h3_perm_test))
)
h3_pass <- all(h3_pvals >= 0.0125, na.rm = TRUE) && (h3_ci_lower < 0 && h3_ci_upper > 0)

# H4: interaction p-value
h4_pass <- anova_h4_art["FrameType:Condition", "Pr(>F)"] < 0.0125

# H5: difference in correlation strength
h5_pass <- h5_perm_p < 0.0125

# H6: difficulty moderation interaction
h6_pass <- anova_h6_art["FrameType:Difficulty", "Pr(>F)"] < 0.0125

# H7: variability effect
h7_pass <- h7_levene_result[["Pr(>F)"]][1] < 0.0125

# H8: encoding strength effect
h8_pass <- h8_reg_p < 0.0125

results_summary <- tibble(
  Hypothesis = c("H1", "H2", "H3", "H4", "H5", "H6", "H7", "H8"),
  Decision = c(
    ifelse(h1_pass, "Pass", "Fail"),
    ifelse(h2_pass, "Pass", "Fail"),
    ifelse(h3_pass, "Pass", "Fail"),
    ifelse(h4_pass, "Pass", "Fail"),
    ifelse(h5_pass, "Pass", "Fail"),
    ifelse(h6_pass, "Pass", "Fail"),
    ifelse(h7_pass, "Pass", "Fail"),
    ifelse(h8_pass, "Pass", "Fail")
  )
)

results_summary
```

# Demographic Hypotheses and Tests

```{r demo-merge}
# Use complete demographic records only
id_source <- if ("Sub ID" %in% names(demo_data_raw)) {
  as.character(demo_data_raw$`Sub ID`)
} else if ("participant_id" %in% names(demo_data_raw)) {
  as.character(demo_data_raw$participant_id)
} else {
  rep(NA_character_, nrow(demo_data_raw))
}

demo_for_summary <- demo_data_raw %>%
  select(-any_of("Condition")) %>%
  mutate(
    id_source = id_source,
    participant_id = str_to_lower(str_extract(id_source, "(?i)sub\\d+")),
    Age = suppressWarnings(as.numeric(Age)),
    Gender = as.character(Gender),
    Handedness = as.character(Handedness),
    Vision = as.character(Vision)
  ) %>%
  left_join(demo_data %>% select(participant_id, Condition), by = "participant_id") %>%
  filter(!is.na(participant_id), participant_id != "")

demo_complete <- demo_for_summary %>%
  filter(!is.na(Age))

participant_demo <- participant_level_desc %>%
  left_join(
    demo_complete %>% select(participant_id, Age, Gender, Handedness, Vision),
    by = "participant_id"
  ) %>%
  filter(!is.na(Age))
```

## D1: Age Balance (NB vs AB)

```{r demo-test-d1-age-balance}
age_NB <- participant_demo %>% filter(Condition == "NB") %>% pull(Age)
age_AB <- participant_demo %>% filter(Condition == "AB") %>% pull(Age)

td1 <- t.test(age_NB, age_AB, var.equal = FALSE)
td1
```

## D2: Gender Balance (NB vs AB)

```{r demo-test-d2-gender-balance}
gender_table <- table(
  participant_demo$Gender,
  participant_demo$Condition
)

td2 <- chisq.test(gender_table)
td2
```

## D3: Age as Moderator (Age x Accuracy)

```{r demo-test-d3-age-accuracy}
td3 <- cor.test(participant_demo$Age, participant_demo$mean_accuracy, method = "pearson")
td3
```

## D4: Gender as Moderator (Accuracy by Gender)

```{r demo-test-d4-gender-accuracy}
acc_female <- participant_demo %>% filter(Gender == "Female") %>% pull(mean_accuracy)
acc_male   <- participant_demo %>% filter(Gender == "Male") %>% pull(mean_accuracy)

td4 <- t.test(acc_female, acc_male, var.equal = FALSE)
td4
```

## Demographic Results Summary (Pass/Fail)

```{r demo-results-summary}
demo_results_summary <- tibble(
  Hypothesis = c("D1", "D2", "D3", "D4"),
  Decision = c(
    ifelse(td1$p.value < 0.05, "Pass", "Fail"),
    ifelse(td2$p.value < 0.05, "Pass", "Fail"),
    ifelse(td3$p.value < 0.05, "Pass", "Fail"),
    ifelse(td4$p.value < 0.05, "Pass", "Fail")
  )
)

demo_results_summary
```
